---
title: "GSE100924"
author: "Diego Vallarino"
date: "8/3/2021"
output:
  html_document:
    toc: yes
    df_print: paged
    toc_float: true
  pdf_document: 
    latex_engine: xelatex
    toc: yes
    number_sections: yes
---
# Exploring microarray data. A naïve (simple) approach

**Summary**	

Zbtb7b is a zinc finger and BTB domain containing transcription factor that activates the thermogenic gene program during brown and beige adipocyte differentiation. Zbtb7b interacts with the long noncoding RNA Blnc1 and hnRNPU to form a ribonucleoprotein transcriptional complex
We used microarray to determine how Zbtb7b regulates brown fat gene expression at ambient room temperature and following cold exposure
 	
**Overall design**	

Wild type and Zbtb7b knockout mice of 10 weeks of age were kept at ambient room temperature (22C) or following cold exposure at 4C for 4 hrs. Brown adipose tissue was harvested for total RNA isolation and microarray analysis.
## Loading the data

```{r, include=FALSE}
library(GEOquery)
gset <- getGEO("GSE100924", GSEMatrix =TRUE, getGPL=FALSE)
if (length(gset) > 1) idx <- grep("GPL17400", attr(gset, "names")) else idx <- 1
gset <- gset[[idx]]
ex <- exprs(gset)
dim(ex)
colnames(ex) <- c(paste("Turner",1:6, sep="_"), paste("Control",1:6, sep="_"))
colnames(ex)
```

## Exploratory analysis with univariate statistics
```{r setup, include=FALSE}
head(ex)
summary(ex)

```

A boxplot of the data shows that values are assymetrically distributed

```{r boxplot1,fig.align='center', fig.cap='',echo=F}
require(knitr)
boxplot(ex, col=c(rep("red", 6) , rep("green", 6)),main="Expression values for\n 6 Turner  and 6 Control samples",
    xlab="Slides",
    ylab="Expression", las=2, cex.axis=0.7, cex.main=0.7)
abline(0,0, col="black")
```

## Data visualization using unsupervised techniques (PCA, Clustering)

Start by computing prncipal components and loadings.
```{r pca}
pcX<-prcomp(t(ex), scale=TRUE) 
loads<- round(pcX$sdev^2/sum(pcX$sdev^2)*100,1)
```

```{r plotPCA, fig=TRUE}
xlab<-c(paste("PC1",loads[1],"%"))
ylab<-c(paste("PC2",loads[2],"%"))
plot(pcX$x[,1:2],xlab=xlab,ylab=ylab)
title("Principal components (PCA)")
text(pcX$x[,1],pcX$x[,2],colnames(ex), pos=4)
```

Alternatively a hierarchichal clustering can be applied to detect any expected (or unexpected grouping of the samples).

```{r codedendrogramcomputeHC}
clust.euclid.average <- hclust(dist(t(ex)),method="ward.D2")
```

```{r plotdendrograms, fig=T}
plot(clust.euclid.average, hang=-1)
```


